|

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.Rmd
update the writing and code in main doc

@@

if (!file.exists(here::here("data/third_study/conv_result_heter.rds"))) {
  conv_result_heter <- list()
  for (i in 1:100)
  {
    conv_result_heter[[i]] <- map(1:2000,
                            function(x) {
                     a <- sample(c(-1, 0, 1), 1)
                     b <- runif(1, 0, 32)
                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
                     x <- switch(x_dist,
                                 uniform = rand_uniform(-1, 1),
                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
                     mod <- heter_model(a = a, b = b, x = x)
                     n <- sample(c(50, 100, 300), 1)
                     tmp_dat <- mod$gen(n)

                     tibble(a = a,
                            b = b,
                            x_dist = x_dist,
                            n = n,
                            effect_size = mod$effect_size(tmp_dat),
                            p_value = mod$test(tmp_dat)$p_value,
                            reject = p_value < 0.05,
                            boot_id = i)
                   }) %>%
  reduce(bind_rows) %>%
  mutate(reject = as.numeric(reject))
  }

@@

  saveRDS(conv_result_heter, here::here("data/third_study/conv_result_heter.rds"))
} else {
  conv_result_heter <- readRDS(here::here("data/third_study/conv_result_heter.rds"))
}

@@

if (!file.exists(here::here("data/third_study/conv_result_heter_SW.rds"))) {
  conv_result_heter_SW <- list()
  for (i in 1:100)
  {
    conv_result_heter_SW[[i]] <- map(1:2000,
                            function(x) {
                     a <- sample(c(-1, 0, 1), 1)
                     b <- runif(1, 0, 32)
                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
                     x <- switch(x_dist,
                                 uniform = rand_uniform(-1, 1),
                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
                                 even_discrete = rand_uniform_d(k = 5, even = TRUE))
                     mod <- heter_model(a = a, b = b, x = x)
                     n <- sample(c(50, 100, 300), 1)
                     tmp_dat <- mod$gen(n)

                     tibble(a = a,
                            b = b,
                            x_dist = x_dist,
                            n = n,
                            effect_size = mod$effect_size(tmp_dat),
                            p_value = shapiro.test(tmp_dat$.resid)$p.value,
                            reject = p_value < 0.05,
                            boot_id = i)
                   }) %>%
  reduce(bind_rows) %>%
  mutate(reject = as.numeric(reject))
  }

@@

  saveRDS(conv_result_heter_SW, here::here("data/third_study/conv_result_heter_SW.rds"))
} else {
  conv_result_heter_SW <- readRDS(here::here("data/third_study/conv_result_heter_SW.rds"))
}

@@

> *"Since all models are wrong the scientist must be alert to what is importantly wrong."* [@REF]

Diagnosing a model is the key to determining whether there is anything importantly wrong. For linear regression analysis, it is typical to interrogate the residuals. Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong. There are many ways that residuals are assessed.

Residuals might be plotted, as a histogram or normal probability plot to examine the distribution. Is it symmetric, and unimodal? Then all is good. If the distribution is skewed or bimodal or multimodal, there are outliers, there is cause for concern. It if also possible to conduct a goodness of fit test, such as a Shapiro-Wilk Normality test [@REF].

It is also recommended to plot residuals against predicted values, and also each of the explanatory variables. If there is no obvious pattern, then all is good. It is a very difficult task for a human judge, though to make a decision that there's nothing there. It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none. Generally, one looks for departures from nothingness like nonlinear dependency or heteroskedasticity. It is also possible to conduct a hypothesis test for non-linear dependence [@RESET], and also use a Breusch-Pagan test [@REF] for heteroskedasticity.

There is an abundance of literature describing appropriate diagnostic methods for linear regression: Cook and Weisberg, ... [@REF]. The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* There is a common guidance by experts that plots are the best for diagnosing model fits.

This is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the commonly used diagnostic plots, the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers. Section XXX describes the experimental setup to enable a comparison between decisopns made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section XXX. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

## Hypothesis testing

Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects. Depending on the alternative hypothesis that is focused on, a variety of tests can be applied. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].

As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers. @montgomery1982introduction suggested that based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies. Not to mention, it is almost impossible to have an exactly correctly specified model in reality. There is a well-known aphorism in statistics stated by George Box - "All models are wrong, but some are useful". This indicates proper hypothesis tests will always reject the null hypothesis as long as the sample size is large enough. The outcome "Not reject" can be interpreted as either "effect size is small" or "sample size is small". and the outcome "reject" doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction. But still, the effectiveness of statistical tests shall not be disrespected. Statistical tests have a chance to provide analysts with unique information. In situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots are needed to be checked, one will have no choice but to fall back on statistical tests if there exists any. A good regression diagnostic practice should be a balanced combination of both methods.

@@

@@

In overall, there were a total of 3200 lineup evaluations made by 160 subjects in experiment I, where 320 lineup evaluations were attention checks and were not used in the following analysis. The collated dataset is provided in `polynomials` of the `visage` `R` package.

@@

polynomials %>%
  group_by(set) %>%
  slice_sample(n = 1) %>%
  ungroup() %>%
  count(pronoun, education, age_group, previous_experience) %>%
  arrange(desc(n)) %>%
  knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"),
               escape = FALSE,
               format = "latex",
               linesep = "",
               booktabs = TRUE, 
               caption  = 'Summary of demographic information', 
               label = "demographic-table" ) %>%
  cat()
```

```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"}
polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(education = fct_relevel(education, "High School or below")) %>%
  ggplot() +
  geom_bar(aes(y = education), orientation = "y") -> education_plot

polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  ggplot() +
  geom_bar(aes(age_group)) -> age_plot

polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>%
  ggplot() +
  geom_bar(aes(pronoun)) -> gender_plot

polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  ggplot() +
  geom_bar(aes(previous_experience)) -> experience_plot

@@

@@

polynomials %>%
  filter(lineup_id == 364)
```

@@

### Shpaes of non-linearity

@@

### Reasons for making selections

@@

```{r cache=TRUE}
heter_result <- read_csv(here::here("data/processed/processed_surveys.csv"))

heter_result <- heter_result %>%
  filter(type != "cubic")

heter_cache <- new.env()

heter_result <- heter_result %>%
  left_join(heter_result %>%
  calc_p_value_multi(detected = "detect", n_sel = "num_selection", cache_env = heter_cache))

heter_result <- heter_result %>%
  mutate(weigthed_detect = ifelse(detect, detect/num_selection, 0)) %>%
  group_by(lineup_id) %>%
  mutate(prop_detect = mean(weigthed_detect))
```

@@

  visual_dat <- heter_result %>%
  mutate(x_dist = ifelse(x_dist == "discrete_uniform", "even_discrete", x_dist)) %>%
  mutate(x_dist = ifelse(x_dist == "neglognormal", "lognormal", x_dist)) %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value, prop_detect), first)) %>%
  filter(ori_difficulty != "extremely_easy") %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  boot_dat(times = 100) %>%
  mutate(cat_b = map_to_group(b))

>>>>

<!-- check the policy -->

@@

if (!file.exists(here::here("data/fourth_study/conv_result_heter.rds"))) {
  conv_result_heter <- list()
  for (i in 1:100)
  {
    conv_result_heter[[i]] <- map(1:2000,
                            function(x) {
                     a <- sample(c(-1, 0, 1), 1)
                     b <- sample(c(0.25, 1, 4, 16, 64), 1)
                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
                     x <- switch(x_dist,
                                 uniform = rand_uniform(-1, 1),
                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
                                 even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
                     mod <- heter_model(a = a, b = b, x = x)
                     n <- sample(c(50, 100, 300), 1)
                     tmp_dat <- mod$gen(n)

                     tibble(a = a,
                            b = b,
                            x_dist = x_dist,
                            n = n,
                            effect_size = mod$effect_size(tmp_dat, type = "kl"),
                            p_value = mod$test(tmp_dat)$p_value,
                            reject = p_value < 0.05,
                            boot_id = i)
                   }) %>%
  reduce(bind_rows) %>%
  mutate(reject = as.numeric(reject))
  }

@@

  saveRDS(conv_result_heter, here::here("data/fourth_study/conv_result_heter.rds"))
} else {
  conv_result_heter <- readRDS(here::here("data/fourth_study/conv_result_heter.rds"))
}

@@

if (!file.exists(here::here("data/fourth_study/conv_result_heter_SW.rds"))) {
  conv_result_heter_SW <- list()
  for (i in 1:100)
  {
    conv_result_heter_SW[[i]] <- map(1:2000,
                            function(x) {
                     a <- sample(c(-1, 0, 1), 1)
                     b <- sample(c(0.25, 1, 4, 16, 64), 1)
                     x_dist <- sample(c("uniform", "normal", "lognormal", "even_discrete"), 1)
                     x <- switch(x_dist,
                                 uniform = rand_uniform(-1, 1),
                                 normal = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))},
                                 lognormal = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))},
                                 even_discrete = rand_uniform_d(-1, 1, k = 5, even = TRUE))
                     mod <- heter_model(a = a, b = b, x = x)
                     n <- sample(c(50, 100, 300), 1)
                     tmp_dat <- mod$gen(n)

                     tibble(a = a,
                            b = b,
                            x_dist = x_dist,
                            n = n,
                            effect_size = mod$effect_size(tmp_dat, type = "kl"),
                            p_value = shapiro.test(tmp_dat$.resid)$p.value,
                            reject = p_value < 0.05,
                            boot_id = i)
                   }) %>%
  reduce(bind_rows) %>%
  mutate(reject = as.numeric(reject))
  }

@@

  saveRDS(conv_result_heter_SW, here::here("data/fourth_study/conv_result_heter_SW.rds"))
} else {
  conv_result_heter_SW <- readRDS(here::here("data/fourth_study/conv_result_heter_SW.rds"))
}

@@

> *"Since all models are wrong the scientist must be alert to what is importantly wrong [@box1976science]."*

Diagnosing a model is the key to determining whether there is anything importantly wrong. For linear regression analysis, it is typical to interrogate the residuals. Residuals summarise what is not captured by the model, and thus provide the capacity to identify what might be wrong. There are many ways that residuals could be assessed.

Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution. Using classical linear regression model as an example, if the distribution is symmetric and unimodal, then it is well-behaved. But if the distribution is skewed, bimodal, multimodal, or even contains outliers, there is cause for concern. It is also possible to conduct a goodness of fit test to examine the distribution of residuals, such as Shapiro-Wilk Normality test [@shapiro1965analysis].

To examine the relationship between residuals, predicted values and each of the explanatory variables, it is recommend to plot them on a scatter plot with residuals on the y-axis. This type of plot is usually called residual plot. In general, one will check whether there is any noticeable pattern to determine if the model is well-behaved. However, it is a very difficult task for a human judge, though to make a decision that there's nothing there. It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none. Generally, one looks for departures from nothingness like nonlinear dependency or heteroskedasticity. It is also possible to conduct a hypothesis test for non-linear dependence [@ramsey_tests_1969], and also use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

There is an abundance of literature describing appropriate diagnostic methods for linear regression: @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. The diligent reader of these sage writings will also notice sentences that express sentiments like *based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies.* There is a common guidance by experts that plots are the best for diagnosing model fits.

This is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the commonly used diagnostic plots, the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers. Section \ref{experimental-design} describes the experimental setup to enable a comparison between decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

<!-- ## Hypothesis testing -->

<!-- Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects. Depending on the alternative hypothesis that is focused on, a variety of tests can be applied. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. -->

<!-- As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers. @montgomery1982introduction suggested that based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies. Not to mention, it is almost impossible to have an exactly correctly specified model in reality. There is a well-known aphorism in statistics stated by George Box - "All models are wrong, but some are useful". This indicates proper hypothesis tests will always reject the null hypothesis as long as the sample size is large enough. The outcome "Not reject" can be interpreted as either "effect size is small" or "sample size is small". and the outcome "reject" doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction. But still, the effectiveness of statistical tests shall not be disrespected. Statistical tests have a chance to provide analysts with unique information. In situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots are needed to be checked, one will have no choice but to fall back on statistical tests if there exists any. A good regression diagnostic practice should be a balanced combination of both methods. -->

@@

Heter Experiment design here

@@

In overall, there were a total of 3200 lineup evaluations made by 160 subjects in both experiment I and experiment II respectively, where 320 lineup evaluations were attention checks and were not used in the following analysis.

The collated dataset is provided in `polynomials` and `heter` of the `visage` `R` package.

@@

polynomials %>%
  mutate(exp = 1) %>%
  bind_rows(mutate(heter, exp = 2)) %>%
  group_by(exp, set) %>%
  slice_sample(n = 1) %>%
  ungroup() %>%
  count(pronoun, education, age_group, previous_experience) %>%
  arrange(desc(n)) %>%
  knitr::kable(col.names = c("Preferred pronoun", "Education", "Age group", "Previous experience", "Subject"),
               escape = FALSE,
               format = "latex",
               linesep = "",
               booktabs = TRUE, 
               caption  = 'Summary of demographic information', 
               label = "demographic-table" ) %>%
  cat()
```

```{r demographic-summary-plot, fig.cap="Summary of demographic information. \\label{fig:demographic-summary-plot}"}
polynomials %>%
  mutate(exp = 1) %>%
  bind_rows(mutate(heter, exp = 2)) %>%
  group_by(exp, set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(education = fct_relevel(education, "High School or below")) %>%
  ggplot() +
  geom_bar(aes(y = education), orientation = "y") -> education_plot

polynomials %>%
  mutate(exp = 1) %>%
  bind_rows(mutate(heter, exp = 2)) %>%
  group_by(exp, set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  ggplot() +
  geom_bar(aes(age_group)) -> age_plot

polynomials %>%
  mutate(exp = 1) %>%
  bind_rows(mutate(heter, exp = 2)) %>%
  group_by(exp, set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>%
  ggplot() +
  geom_bar(aes(pronoun)) -> gender_plot

polynomials %>%
  mutate(exp = 1) %>%
  bind_rows(mutate(heter, exp = 2)) %>%
  group_by(exp, set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  ggplot() +
  geom_bar(aes(previous_experience)) -> experience_plot

@@

### Experiment I

@@

polynomials %>%
  filter(lineup_id == 364)

polynomials %>%
  filter(e_sigma == 4) %>%
  group_by(lineup_id) %>%
  summarise(p_value = first(p_value)) %>%
  mutate(reject = p_value < 0.05) %>%
  {mean(.$reject)}

4 * 4 * 3 * 5
```

@@

#### Shpaes of non-linearity

@@

#### Reasons for making selections

@@

### Experiment II

```{r eval = FALSE}
heter %>%
  filter(b < 3) %>%
  group_by(lineup_id) %>%
  summarise(p_value = first(p_value), n_eval = n()) %>%
  filter(n_eval > 3) %>%
  mutate(reject = p_value < 0.05) %>%
  {mean(.$reject)}

@@

  visual_dat <- heter %>%
  mutate(x_dist = ifelse(x_dist == "discrete_uniform", "even_discrete", x_dist)) %>%
  mutate(x_dist = ifelse(x_dist == "neglognormal", "lognormal", x_dist)) %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value, prop_detect), first)) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  boot_dat(times = 100) %>%
  mutate(cat_b = map_to_group(b))
